---
title: "P8105 Homework 2"
output: github_document
author: Shritama Ray
date: 10/5/2022
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## Problem 1: NYC Transit Data
*My attempt to solve without the solutions*

Read & Clean the Dataset
```{r read and clean}
transit_df = read_csv("./Data/subway_data.csv") %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada)%>%
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE)) 
```

**Summary:**
This dataset contains information on the NYC subway system. Specifically, the dataset describes the routes and locations of the stations, along with some other features such as ADA accessibility, vending machines, and entrance type. So far I've manipulated the data to create clean variable names, remove the variables I am not interested in, and convert the entry values to logical operators. The resulting dataset has **1868 rows** and **19 columns.** The data is **NOT tidy.** The route variables are in a wide format, resulting in many NA values. 

**Questions:**

1. How many distinct stations are there?
```{r distinct stations}
nrow(distinct(transit_df, line, station_name))
```
There are **465** distinct stations.

2. How many stations are ADA compliant?
```{r ada compliant}
distinct(transit_df, line, station_name, ada) %>%
  filter(ada == TRUE) %>%
  nrow()
```
**84** stations are ADA compliant

3. What proportion of station entrances/exits without vending allow entrance?
```{r no vending, allow entrance}
filter(transit_df, vending == "NO" & entry == TRUE) %>%
  nrow()

filter(transit_df, vending == "NO") %>%
  nrow()
```
**69/183** of the entrances/exits without vending allow entry

Reformat the data so that route number and route name are distinct variables:
```{r reformat long}
  transit_tidy = 
  mutate_at(transit_df, c('route8','route9','route10','route11'), as.character) %>%
  pivot_longer(route1:route11, 
               names_to = "route_number", 
               names_prefix = "route", 
               values_to = "route_name")
```

How many distinct stations serve the A train?
```{r distinct stations A train}
nrow(
  filter(
    distinct(transit_tidy, line, station_name, route_name),route_name == "A"))
```
**60** distinct stations serve the A train.

Of the stations, that serve the A train, how many are ADA complaint?
```{r ada compliant A train}
nrow(
  filter(
    distinct(transit_tidy, line, station_name, route_name, ada),
    route_name == "A" & ada == TRUE))
```
**17** of the stations that serve the A train are ADA compliant. 

## Problem 2: Mr. Trash Wheel 

Read and clean the Mr. Trash Wheel dataset:
```{r read and clean trash_wheel}
library(readxl)
mr_trash = read_excel("./Data/trash_wheel.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N549") %>%
  janitor::clean_names() %>%
   filter(dumpster != '') %>%
   mutate(sports_balls = as.integer(sports_balls), year = as.integer(year))
```

Read and clean the Professor Trash Wheel sheet:
```{r read & clean professor trash wheel}
prof_trash = read_excel("./Data/trash_wheel.xlsx", sheet = "Professor Trash Wheel", range = "A2:M96") %>%
  janitor::clean_names() %>%
   filter(dumpster != '')
```

Combine the two datasets:
```{r combine trash wheels}
#First create a new variable to track each dataset
mr_trash = mutate(mr_trash, wheel = "Mr.")
prof_trash = mutate(prof_trash, wheel = "Prof")

#combine
combined_trash = bind_rows(mr_trash, prof_trash)
```

This combined dataset has a total of **`r nrow(combined_trash)`** rows and **`r ncol(combined_trash)`** columns. Some of the key variables include the dumpster, month/year of collection, weight in tons, # of homes powered, and the count of types of trash, such as grocery bags, cigarette butts, etc. 

What was the total weight of trash collected by Professor Trash Wheel?

Professor Trash Wheel collected a total of **`r sum(pull(filter(combined_trash, wheel == "Prof"), weight_tons))`** tons of trash. 

What was the total number of sports balls collected by Mr. Trash Wheel in 2020?

Mr. Trash Wheel collected a total of **`r sum(pull(filter(combined_trash, wheel == "Mr." & year == 2020), sports_balls))`** sports balls in 2020. 

## Problem 3: FiveThirtyEight

Read & Clean pols-month.csv:
```{r pols read & clean}
pols_month = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(month = month.name[as.integer(month)],
         president = recode(prez_dem, '1' = "dem", '0' = "gop")) %>%
  select(-day, -prez_dem, -prez_gop)
```

Read & clean snp.csv:
```{r read & clean snp}
snp = read_csv("./data/fivethirtyeight_datasets/snp.csv") 

#Convert year to 4 digits
four_year = as.Date(snp$date,"%m/%d/%y")
four_year = as.Date(ifelse(four_year > Sys.Date(), format(four_year, "19%y-%m-%d"), format(four_year)))
snp[,"date"] <- four_year

#Clean data
snp = 
  separate(snp, date, into = c("year", "month", "day"))%>%
  mutate(month = month.name[as.integer(month)]) %>%
  select(year, month, close)
```

Read & clean unemployment dataset:
```{r unemployment}
unemployment = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(Jan:Dec,
               names_to = "month",
               values_to = "percent") %>%
  mutate(year = as.character(Year), 
         month = month.name[match(pull(.,month),month.abb)]) %>%
  arrange(year, month)

#Note: I removed the janitor::clean_names() statement since it made the months lower case
```


Merge the Datasets:
```{r 538 merge}
merged_538 = left_join(pols_month, snp, by = c("year","month")) %>%
              left_join(unemployment, by = c("year","month"))
```

**Summary:**
These datasets contain various information about politics and economics. The pols_month dataset contains info on national politicians and their party at various points of time. The snp dataset contains the closing values of the S&P stock market index at various points of time. The unemployment dataset has the monthly national unemployment rate over several years. The merged dataset contains **`r nrow(merged_538)`** rows and **`r ncol(merged_538)`** columns. The data contains information over the years **`r min(merged_538$year)`** to **`r max(merged_538$year)`.** It contains information on the number of politicians from each party, the S&P closing value, and unemployment rate at each month over this period. This dataset can be used to analyze associations between political landscapes and economic success. 
